---
title: "Table Visualization Part 1"
subtitle: "R Shiny with Loan"
author: |
  | [Youtube: https://www.youtube.com/@rshiny_with_loan](https://www.youtube.com/@rshiny_with_loan)
  | [Website: https://loankimrobinson.com/](https://loankimrobinson.com/)
  | [Github: https://github.com/loankimrobinson/rshiny_with_Loan](https://github.com/loankimrobinson/rshiny_with_Loan)
  | [LinkedIn: https://www.linkedin.com/in/loan-robinson-86776994/](https://www.linkedin.com/in/loan-robinson-86776994/)
date: "`r Sys.Date()`"
indent: TRUE
output:
  html_document: 
    self_contained: false
    toc: true
    toc_depth: 4
    toc_float: false
    theme: cerulean
    highlight: zenburn
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(reactable.static = TRUE)
library(reactable)
library(reactablefmtr)
library(htmltools)
library(dplyr)
library(fontawesome)
library(DT)
```


```{css ref.label="styles", echo=FALSE}

```


```{css ref.label="pkgdown_styles", echo=FALSE}
```


### Features

  * `Sticky` Columns
  * Add logo 
  * Add `data_bars()`
  * Add hyperlink
  * Nested `reactable`
  * `icon_assign()`
  * `htmltools()`
  * theme
  * Cell Style

### R Library

  * reactable
  * reactablefmtr
  * dplyr
  * htmltools
  * DT
  
### What I should think before starting building a Table Visualization

 
#### 1. Idea

 

  * What do I want to present data?

  * Any summary data I should include?

  * What kind of charts I should use?

  * Style or No Style for each variable.

  * Which variables are the most important?

 

#### 2. What Are the Most Important Factors When Applying the `reactable` Package.

 
  * Size of data (If it's too heavy, use filter options, try to use `JS`)

  * Overwhelming with many charts.

  * Overwhelming with many colors.

  * Variable position is important.

  * The width of column is important too.

  * Start building a visualization table column by column

 

#### 3. Let's start making a beatiful, presentable table


#####     a. Without Applying Reactable Package

```{r DT-table, echo=FALSE, warning=FALSE, message=FALSE}

data <- MASS::Cars93
data <- data[data$Make != "Mazda RX-7", ]
data$DriveTrain <- ifelse(data$DriveTrain  == "Front", "FWD",
                          ifelse(data$DriveTrain  == "Rear", "RWD","4WD"))
data$Manufacturer[data$Manufacturer == "Chrylser"] <- "Chrysler"
data$Manufacturer <- as.character(data$Manufacturer )
data$Make <- as.character(data$Make)
data$Cylinders <- as.numeric(data$Cylinders)
summary_dt <- data %>% group_by(Manufacturer) %>% arrange(desc(Max.Price)) %>% slice(1)
summary_dt <- summary_dt[, c("Manufacturer","Make","Price",
                             "MPG.city", "MPG.highway", "Type", 
                             "EngineSize", "Horsepower","Cylinders",
                             "Length", "Width", 
                             "Weight")]
summary_dt$type_col <- ifelse(summary_dt$Type == "Small",'#011936',
                              ifelse(summary_dt$Type == "Midsize",'#465362',
                                     ifelse(summary_dt$Type == "Large",'#82A3A1',
                                            ifelse(summary_dt$Type =="Van",'#9FC490','#C0DFA1'))))
data <- data[,c("Manufacturer", "Make","Price","MPG.city","MPG.highway","Horsepower")]

summary_dt <- data.frame(summary_dt)
DT::datatable(
              options = list(pageLength = 5, autoWidth = TRUE),
              summary_dt[,colnames(summary_dt) != "type_col"])
```

</br>

#####     b. With Reactable Package

```{r table, echo=FALSE, warning=FALSE, message=FALSE}

tbl <- reactable(
  summary_dt,
  width = 1200,
  pagination = TRUE,
  defaultPageSize = 10,
  defaultSorted = "Price",
  defaultSortOrder = "desc",
  defaultColDef = colDef(
    vAlign = "center",
    headerVAlign = "bottom",
    class = "cell",
    headerClass = "header"
  ),
  detail = function(index){
    index <- summary_dt[index,"Manufacturer"]
    dt_detail <- data[data$Manufacturer %in% index,]
    htmltools::div(style = "padding-left:100px;padding-top:60px;padding-bottom:60px;background-color:#FCFCFC;",
                   reactable(
                     dt_detail,
                     width = "800",
                     defaultSorted = "Price",
                     defaultSortOrder = "desc",
                     defaultColDef = colDef(
                       headerStyle = list(background = "#f7f7f8", color="gray"),
                       vAlign = "center",
                       headerVAlign = "bottom",
                       class = "cell",
                       headerClass = "header"
                     ),
                     theme = clean(),
                     columns = list(
                       Manufacturer = colDef(show = F),
                       Make  = colDef(html = TRUE, align = "center", minWidth = 150,cell = JS('
                        function(cellInfo) {
                        // Render as a link
                          const url = `https://wikipedia.org/wiki/${cellInfo.value}`
                          return `<a href="${url}" target="_blank">${cellInfo.value}</a>`
                          }
                        ')),
                       Price = colDef(align = "center", minWidth = 150 ),

                       MPG.city= colDef(
                         "MPG in City",
                         minWidth = 120,
                         align = "left",
                         cell = data_bars(
                           data = dt_detail,
                           min_value = min(dt_detail$MPG.city)-min(dt_detail$MPG.city)/15,
                           max_value = max(dt_detail$MPG.city)+max(dt_detail$MPG.city)/15,
                           fill_color = 'slategrey',
                           background = '#FFFFFF',
                           bar_height = 7,
                           
                           text_position = 'outside-end',
                           icon = 'circle',
                           icon_color = 'slategrey',
                           icon_size = 15,
                           text_color = 'slategrey',
                           round_edges = TRUE
                         )),
                       
                       MPG.highway = colDef(
                         "MPG in Highway",
                         align = "left",
                         minWidth = 120,
                         cell = data_bars(
                           data = dt_detail,
                           min_value = min(dt_detail$MPG.highway)-min(dt_detail$MPG.highway)/15,
                           max_value = max(dt_detail$MPG.highway)+max(dt_detail$MPG.highway)/15,
                           fill_color = 'slategrey',
                           background = '#FFFFFF',
                           bar_height = 7,
                           
                           text_position = 'outside-end',
                           icon = 'circle',
                           icon_color = 'slategrey',
                           icon_size = 15,
                           text_color = 'slategrey',
                           round_edges = TRUE
                         )
                       ),
                                              
                       Horsepower = colDef(
                         "Horse Power",
                         align = 'center',
                         minWidth = 100,
                         style = function(value) {
                           if (value > 200) {
                             color <- "#008000"
                             fontWeight = "bold"
                           } else {
                             color <- "gray"
                             fontWeight = "normal"
                           }
                           list(color = color, fontWeight = fontWeight)
                         }
                       )
                     )
                   )
                  )
  },
  columns = list(
    Make = colDef(show = FALSE),
    
    Manufacturer = colDef(
      "Make",
      minWidth = 250,
      sticky = "left",
      style = list(borderRight = "1px solid #eee",backgroundColor = "#FCFCFC"),
      cell = function(value, index) {
        div(
          class = "manufacturer",
          span(img(class = "manufacturer-logo", alt = paste(value, "logo"), src =sprintf("car_logo/%s.png", value))),
          span(class = "manufacturer-name", sprintf("%s ", summary_dt[index, "Make"]))
        )
      }
    ),
    
    Price = colDef(
      "Price",
      align = 'left',
      minWidth = 150,
      cell = data_bars(
        data = summary_dt,
        fill_color = '#EEEEEE',
        number_fmt = scales::number_format(accuracy = 0.1),
        text_position = 'outside-end',
        max_value = max(summary_dt$Price) + max(summary_dt$Price)/5,
        icon = 'circle',
        icon_color = 'slategrey',
        icon_size = 15,
        #text_color = 'slategrey',
        round_edges = TRUE
      )
    ),
    
    MPG.city = colDef(
      "MPG in City",
      align = 'center',
      minWidth = 100
    ),
    
    MPG.highway = colDef(
      "MPG in Highway",
      align = 'center',
      minWidth = 100
    ),
    
    Type = colDef(
      minWidth = 120,
      align = 'center',
      cell = color_tiles(
        data = summary_dt,
        color_ref = 'type_col',
        opacity = 0.7,
        bold_text = FALSE,
        box_shadow = TRUE
      )
    ),
    
    EngineSize = colDef(
      "Engine Size",
      align = 'center',
      minWidth = 100
    ),
    
    Cylinders = colDef(
      "Cylinders",
      align = 'center',
      minWidth = 120,
      cell = icon_assign(
        align = 'center',
        summary_dt,
        fill_color = 'slategrey',
        empty_color = 'lightgrey',
        empty_opacity = 0.8,
        icon_size = 12
      )
    ),
    
    Horsepower = colDef(
      "Horse Power",
      align = 'center',
      minWidth = 100,
      style = function(value) {
        if (value > 200) {
          color <- "#008000"
          fontWeight = "bold"
        } else {
          color <- "gray"
          fontWeight = "normal"
        }
        list(color = color, fontWeight = fontWeight)
      }
    ),
    
    type_col = colDef(show = FALSE)
  )
)

tbl


```




```{css styles, eval=FALSE, include = FALSE}

.author {font-size: 15px;font-family: "American Typewriter", serif;color:black;}
.title {font-family: "Snell Roundhand", cursive; font-size: 35px;font-weight:bold;}
.date {font-family: "American Typewriter", serif; font-size: 20px;font-weight:bold;}
.subtitle {font-family: "American Typewriter", serif; font-size: 20px;font-weight:bold;}
h3, h4 {font-family: "American Typewriter", serif; font-size: 20px;font-weight:bold;}
p {font-family: "American Typewriter", serif;}

.header {
  border-bottom-color: #555;
  font-size: 1.5rem;
  font-weight: 700;
  /*text-transform: uppercase;*/
  color:#465362; 
}
/* Highlight headers when sorting */
.header:hover,
.header:focus,
.header[aria-sort="ascending"],
.header[aria-sort="descending"] {
  background-color: #eee;
}

.manufacturer {
  display: flex;
  align-items: center;
}
.manufacturer-logo {
  height: 2rem;
  width: 4rem;
  border: 1px solid #f0f0f0;
}
.manufacturer-name {
  margin-left: 1rem;
  font-size: 1.4rem;
  font-weight: 700;
  color:#465362; 
}
.manufacturer-origin {
  margin-left: 0.35rem;
  color: hsl(0, 0%, 45%);
  font-size: 1.125rem;
  font-weight: 700;
}

a {color:#9F2B68;font-weight:normal}

```


```{css pkgdown_styles, include=FALSE}

/* pkgdown articles */
.row > main {
  max-width: 1800px;
}
.page-header {
  display: none;
}
/* rmarkdown html documents */
.main-container {
  max-width: 1600px !important;
  margin-left: 80px;
  margin-right: 80px;
  color:#184558;
  font-family: "American Typewriter", serif; 
  font-size: 15px;
}
```


</br>



